A Theory for Distributed Database Transactions

### 2019-06-12 2018_Transactions/articleTransactions.tex
Simplified the definition of history to one relation, prec.

\section{Traces}
\label{sct:Traces}
	We need the notion of {\em event}\footnote{Cambridge Dictionary defines ``event'' as: anything that happens} to describe things that happen in a distributed information system.
	Being distributed means that we must assume that the temporal order of events is partial:
	We do not necessarily know in which order two events have happened.
	To describe this, we define traces algebraically corresponding to Foster~\cite{Foster17b}.

	We define traces as a set ${\mathcal T}$ with
	a concatenation operator $+:\mathcal T\rightarrow\mathcal T\rightarrow\mathcal T$ and
	the empty trace $0:\mathcal T$, which obey the following axioms.
\begin{definition}[Trace Algebra]
\label{def:Trace Algebra}
\item   A trace algebra $(\mathcal T, +, 0)$, is an algebra that satisfies: 
\begin{eqnarray}
	x+(y+z)&=&(x+y)+z\label{axi:associative}\\
	0+x&=&x+0\label{axi:lr-identity}\\
	x+y\ =\ x+z&\Rightarrow&y=z\label{axi:right cancellation}\\
	x+z\ =\ y+z&\Rightarrow&x=y\label{axi:left cancellation}\\
	x+z\ =\ 0&\Rightarrow&x=0\label{axi:no inverse}
\end{eqnarray}
\end{definition}
	In terms of universal algebra, equations~\ref{axi:associative} and~\ref{axi:lr-identity} state that a trace algebra is a {\em monoid}.
	Equations~\ref{axi:right cancellation} and~\ref{axi:left cancellation} state that $+$ is {\em cancellative}.
	Equation~\ref{axi:no inverse} makes the cancellative monoid into a trace algebra.
	It implies that there is no inverse.

	In terms of information systems, a trace represents history.
	For that purpose it makes sense to define the notion of prefix, so we can define notions like shared history.
\begin{definition}[Prefix]
\label{def:Prefix}
\begin{eqnarray}
	x\leq y&\Rightarrow&\exists z:\ x+z=y
\end{eqnarray}
\end{definition}
	We also define subtraction of traces as a way to "cut off" a prefix from a trace.
\begin{definition}[Subtract]
\label{def:Subtract}
\label{def:Subtract}
\begin{array}[t]{rcll}
	y-x&=&\text{if }x\leq y:&\exists z: y=x+z\Rightarrow z\\
	   & &\text{otherwise}:&0
\end{array}
\end{definition}

\section{History}
\label{sct:History}
	The history of a distributed information system (DIS) consists of events and an ordering on these events.
	Because a DIS is distributed,
	we may not know the temporal order of two events in the history structure.
	For this reason we describe the history structure of a DIS as a partial order of events.

	We use variables $e$, $e_1$, and $e_2$ to represent events.
	Let $\mathbb E$ represent the set of all conceivable events.
	We also introduce the notion of {\em history structure}
	to represent all events that have happened up to some point in time and the precedences between those events.
	In this paper, variables $h$, $h'$, $h_1$, and $h_2$ represent history structures.
	Let $\mathbb H$ be the set of all conceivable history structures.

	A {\em history structure} is a tuple $\la \Events, \id{prec}\ra$ in which
\begin{eqnarray}
	&&\Events\ \subseteq\ \mathbb E\\
	&&\id{prec}\ \subseteq\ \Events\times\Events\label{req:precInE}\\
	&&(e,e)\notin\id{prec}\label{req:prec irreflexive}\\
	&&(e_1,e_2)\in\id{prec}^+\ \wedge\ (e_2,e_1)\in\id{prec}^+\ \Rightarrow\ e_1=e_2\label{req:precPlus asy}
\end{eqnarray}
	In a history structure $\la \Events, \id{prec}\ra$ set $\Events$ represents the events that have been registered up to some point in time.
	For every event $e\in\Events$ we can say that it has happened (at some time in the past).
	Let $e_1\ \id{prec}\ e_2$ represent the fact that event $e_1$ has happened before event $e_2$,
	making $\id{prec}$ a relation on events in $\Events$.
	Requirement~\ref{req:prec irreflexive} excludes the situation that $e$ has happened before itself.
	To represent time faithfully, we require that $\la\Events,\id{prec}^*\ra$ is a partially ordered set,
	meaning that $\id{prec}^*$ is reflexive, transitive, and antisymmetric.
	For this reason we have requirement~\ref{req:precPlus asy}, which says that the transitive closure $\id{prec}^+$ must be antisymmetric.

	In a history structure $\la \Events, \id{prec}\ra$, an event $e$ that precedes another event is called a {\em past} event,
	which we define as \ $\exists e_1\in\Events: e\ \id{prec}\ e_1$.
	All other events are called {\em current} events in that history structure.
	Function $\id{current}:\mathbb H\rightarrow\powerset{\Events}$ computes the current events in a history structure.
	Events in $\Events$ that have no preceding events are called {\em initial} events.
	Function $\id{past}$ computes the past events and $\id{initial}$ computes the initial events.
\begin{eqnarray}
	\id{current}(\la \Events, \id{prec}\ra)&=&\{ c\in\Events|\ \neg\exists e_1\in\Events: e\ \id{prec}\ e_1 \}\\
	\id{initial}(\la \Events, \id{prec}\ra)&=&\{ c\in\Events|\ \neg\exists e_1\in\Events: e_1\ \id{prec}\ e \}
\end{eqnarray}
	We define the {\em closure} of a history structure as the history structure whose order is the reflexive, transitive closure of the original ordering relation.
\begin{eqnarray}
	\la \Events, \id{prec}\ra^*&=&\la \Events, \id{prec}^*\ra\label{def:closure}
\end{eqnarray}
	Equation~\ref{def:closure} allows us to write $h^*$ for elements of $\mathbb H$.
	The order of $h^*$ is reflexive, transitive, and antisymmetric.

	The special case in which $\la\Events,\id{prec}^*\ra$ is a total order is called a {\em sequential history structure}.
	Note that any nonempty finite sequential history structure has one current event and one first event.
	Other finite history structures may have more.
	In a sequential history structure, every event except the first one has precisely one preceding event.
	In general, however, events may have multiple preceding events and multiple succeeding events.

	We extend the notion of subset to historic structures and use the subset symbol $\subseteq$ for that purpose.
	We use $h\subseteq h'$ to state that $h$ lies in the past of $h'$.
	Let $h=\la \Events, \id{prec}\ra$ and $h'=\la \Events', \id{prec}'\ra$ in the definition of $h\subseteq h'$:
\begin{equation}
\label{def:lies in the past}
	h\subseteq h'\ \Leftrightarrow\ 
	\Events\subseteq\Events'\ \wedge\ \id{initial}(h)\subseteq\id{initial}(h')\ \wedge\ \id{prec}^*\subseteq\id{prec}'^*
\end{equation}

>>>>

\section{History}
\label{sct:History}
	A treatment of time lays the foundation of our theory of distributed information systems (DIS).
	Because of the distributed nature of a DIS,
	we may not know in which order two events\footnote{Cambridge Dictionary defines ``event'' as: anything that happens} have happened.
	For this reason we describe the history of a DIS as a partially ordered relation of events.

	We use variables $e$, $e_1$, and $e_2$ to represent events.
	Let $\mathbb E$ represent the set of all conceivable events.
%	We also introduce the notion of {\em history structure}
%	to represent all events that have happened up to some point in time and the precedences between those events.
%	In this paper, variables $h$, $h'$, $h_1$, and $h_2$ represent history structures.

	We represent the ``history of a DIS up to some point in time'' by a relation $\id{prec}$ (pronounce: {\em precedes}),
	being a set of pairs from $\mathbb E$.
\[\id{prec}\subseteq\mathbb E\times\mathbb E\]
	Let $e_1\ \id{prec}\ e_2$ represent the fact that event $e_1$ has happened before event $e_2$,
	making $\id{prec}$ a relation on events in $\mathbb E$.
	We require:
\begin{eqnarray}
	&&(e,e)\notin\id{prec}\label{req:prec irreflexive}\\
	&&(e_1,e_2)\in\id{prec}^+\ \wedge\ (e_2,e_1)\in\id{prec}^+\ \Rightarrow\ e_1=e_2\label{req:precPlus asy}
\end{eqnarray}
	The events in the DIS, of which $\id{prec}$ represents the history, are 
\begin{eqnarray}
	&&\id{events}(\id{prec})\ =\ \{e_1|(e_1,e_2)\in\id{prec}\}\cup\{e_2|(e_1,e_2)\in\id{prec}\}\label{req:events}
\end{eqnarray}
	Requirement~\ref{req:prec irreflexive} excludes the situation that $e$ has happened before itself.
	To represent time faithfully, equation~\ref{req:precPlus asy} requires that the transitive closure $\id{prec}^+$ is antisymmetric.
	This makes $\id{prec}^*$ a reflexive, transitive, and antisymmetric relation, which lets us use it as a temporal order between events.
	So we shall call $\id{prec}$ "a history" because it represents all events that have happened up to a certain point in time and the (partial) order between them.
	Let $\mathbb H$ be the set of all conceivable histories, meaning that every $\id{prec}\in\mathbb H$.

	In a history $\id{prec}$, an event $e$ that precedes another event is called a {\em past} event,
	which we define as \ $\exists e_1:\ e\ \id{prec}\ e_1$.
	All other events are called {\em current} events in $\id{prec}$.
	Function $\id{current}$ computes the current events in history $\id{prec}$.
	Function $\id{initial}$ computes the initial events, which are the events without preceding events.
\begin{eqnarray}
	\id{current}(\id{prec})&=&\{e |\ \exists e_1: (e_1,e)\in\id{prec}\ \wedge\ \neg\exists e_2: (e,e_2)\in\id{prec}\}\\
	\id{initial}(\id{prec})&=&\{e |\ \exists e_1: (e,e_1)\in\id{prec}\ \wedge\ \neg\exists e_2: (e_2,e)\in\id{prec}\}
\end{eqnarray}

	The special case in which $\id{prec}^*$ is a total order is called a {\em sequential history}.
	Note that any nonempty finite sequential history has one current event and one initial event.
	Other finite histories may have more.
	In a sequential history structure, every event except the initial one has precisely one preceding event.
	In general, however, events may have multiple preceding events and multiple succeeding events.

	We define $\leq$ to represent the fact that history $\id{prec}$ has preceded history $\id{prec}'$ in time.
	It is defined by:
\begin{equation}
\label{def:lies in the past}
	\id{prec}\leq\id{prec}'\ \Leftrightarrow\ \id{prec}^*\subseteq\id{prec}'^*
\end{equation}

\section{Traces}
\label{sct:Traces}
	A transaction in a DIS may be perceived as a distributed action,
	which leaves a trace of events in the history of a DIS.
	Let $\mathbb T$ be the set of all conceivable traces.
	With \id{prec} being a history, we define that every pair $(e_1,e_2)\in\id{prec}^*$ is a trace.
	So the traces of history $\id{prec}$ are denoted ${\mathcal T}_{\id{prec}}$ with ${\mathcal T}_{\id{prec}}\subseteq\mathbb T$.
\begin{definition}[Traces]
\label{def:Traces}
\item   The set of traces of history $\text{prec}$ is defined by ${\mathcal T}_{\text{prec}}\ =\ \text{prec}^*$.
\end{definition}
	Pair $(e,e)$ represents the empty trace for any $e\in\id{events}(\id{prec})$ because $(e,e)\in\id{prec}^*$.
	We will denote the empty trace as $0$ and define $(e_1,e_1)=0=(e_2,e_2)$.
	If $(e_1,e)\in\id{prec}^*$ and $(e,e_2)\in\id{prec}'^*$ then
	we define the concatenation operator $+:\mathbb T\rightarrow\mathbb T\rightarrow\mathbb T$ by
\begin{definition}[Concatenation of traces]
\label{def:concatenation}
\item   $(e_1,e)+(e,e_2)=(e_1,e_2)$
\end{definition}
	Note that the definition of $+$ is partial, because $(e_1,e_2)+(e_3,e_4)$ is not defined unless $e_2=e_3$.
	Also, $(e_1,e)+(e,e_2)$ is defined in the combined history $\id{prec}^*\cup\id{prec}'^*$
	because $(e_1,e_2)\in(\id{prec}\cup\id{prec}')^*$.

	This definition of trace is an interpretation of trace algebra~\cite{Foster17b}, which describes discrete-event systems.
	Other interpretations of this algebra comprise real-time systems and hybrid systems, but that is beyond the scope of this paper.

	The concatenation operator $+$ and empty trace $0$ satisfy the axioms of trace algebras: (proof needed)
\begin{eqnarray}
	x+(y+z)&=&(x+y)+z\label{axi:associative}\\
	0+x&=&x+0\label{axi:lr-identity}\\
	x+y\ =\ x+z&\Rightarrow&y=z\label{axi:left cancellation}\\
	x+z\ =\ y+z&\Rightarrow&x=y\label{axi:right cancellation}\\
	x+z\ =\ 0&\Rightarrow&x=0\label{axi:no inverse}
\end{eqnarray}
	In terms of universal algebra, equations~\ref{axi:associative} and~\ref{axi:lr-identity} state that a trace algebra is a {\em monoid}.
	Equations~\ref{axi:left cancellation} and~\ref{axi:right cancellation} state that $+$ is {\em cancellative}.
	Equation~\ref{axi:no inverse} makes the cancellative monoid into a trace algebra.
	It implies that there is no inverse.

	It makes sense to define the notion of prefix, so we can define notions like shared history.
\begin{definition}[Prefix]
\label{def:Prefix}
\begin{eqnarray}
	x\leq y&\Rightarrow&\exists z:\ x+z=y
\end{eqnarray}
\end{definition}
	We also define subtraction of traces as a way to "cut off" a prefix from a trace.
\begin{definition}[Subtract]
\label{def:Subtract}
\begin{array}[t]{rcll}
	y-x&=&\text{if }x\leq y:&\exists z: y=x+z\Rightarrow z\\
	   & &\text{otherwise}:&0
\end{array}
\end{definition}
